### 2016-05-25 analysis/efron.fcr.Rmd
add examples

  res.ash = ashr::ash(z,1,mixcompdist="halfuniform")
  CImatrix= ashr::ashci(res.ash,level=0.95)
```

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for the "significant" observations with Bayes rule (green), similar to Figure 8 from Efron. Note that the lower 97.5% point is pretty accurate, but the upper 97.5% point is curtailed - presumably due, at least in part, to the short tails of the uniform mixture.
```{r}
  BayesComparePlot=function(CImatrix,altmean=-3,...){
    plot(z,mu,xlim=c(-8,0),...)
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

>>>>

  res.ash = ashr::ash(z,1,mixcompdist="halfuniform")

```

We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr etc. 
```{r}
  true.g = ashr::normalmix(c(0.9,0.1),c(0,-3),c(0,1))
  res.ash.true = ashr::ash(z,1,g=true.g,fixg=TRUE)
```

Here we can see how the partition of $z$ scores compares with the truth. Note the effect of the unimodal assumption is to extend the inferred alternative distribution toward 0.
```{r}
  source("../R/nullalthist.R")
  par(mfcol=c(2,1))
  nullalthist(z,lfdr=res.ash.true$lfdr,main="true partition")
  nullalthist(z,lfdr=res.ash$lfdr,main="inferred partition")
```

Comparing the inferred Posterior Means, lfdr, and lfsr with the true values of these quantities, we find reassuringly good correspondence.
```{r}
  par(mfcol=c(1,1))
  plot(res.ash.true$PosteriorMean,res.ash$PosteriorMean,xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
  abline(a=0,b=1,col=2,lwd=3)

    plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
  abline(a=0,b=1,col=2,lwd=3)

  plot(res.ash.true$lfsr,res.ash$lfsr,xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
  abline(a=0,b=1,col=2,lwd=3)

```

We can also run qvalue and locfdr. We see that locfdr perhaps performs a bit better than ash for the decomposition here, but the estimated local fdrs are pretty similar. Here qvalue does less well because of the asymmetry which we didn't take account of.
```{r}
library(qvalue)
library(locfdr)
res.locfdr = locfdr(z,nulltype = 0)
res.qvalue = qvalue(p=pchisq(z^2,df=1,lower.tail = FALSE))  
#res.pos.qvalue = qvalue(p=pchisq(z[z>0]^2,df=1,lower.tail = FALSE))  
#res.neg.qvalue = qvalue(p=pchisq(z[z<0]^2,df=1,lower.tail = FALSE))  
#df1= data.frame(Truth = res.ash.true$lfdr,Estimate=res.locfdr$fdr,Method="locfdr")
#df2= data.frame(Truth = res.ash.true$lfdr,Estimate=res.ash$lfdr,Method="ash.hu")
#df3 =data.frame(Truth = res.ash.true$lfdr,Estimate=res.qvalue$lfdr,Method="qvalue")
#df4 =data.frame(Truth = res.ash.true$lfdr[z<0],Estimate=res.neg.qvalue$lfdr,Method="qvalue (-)")
#df5 =data.frame(Truth = res.ash.true$lfdr[z>0],Estimate=res.pos.qvalue$lfdr,Method="qvalue (+)")

#df = rbind(df1,df2,df3,df4,df5)
#ggplot2::qplot(data=df,x=Truth,y=Estimate,facets= .~Method)

par(mfrow=c(1,3))
plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)
plot(res.ash.true$lfdr,res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)
plot(res.ash.true$lfdr,res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1,col=2,lwd=3)

```

# Uniform tail curtails Credible Intervals

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for the "significant" observations with Bayes rule (green), similar to Figure 8 from Efron. Note that the lower 97.5% point is pretty accurate, but the upper 97.5% point is curtailed - presumably due, at least in part, to the short tails of the uniform mixture.
```{r}
  CImatrix= ashr::ashci(res.ash,level=0.95)
  BayesComparePlot=function(CImatrix,altmean=-3,...){
    plot(z,mu,xlim=c(-8,0),...)
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

### 2016-05-25 analysis/index.Rmd
add examples

*  [Illustrate Zero Assumption](makefig_FDReg.html)
*  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
*  [Adaptive Shrinkage](adaptive_shrinkage.html)
*  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
*  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)

>>>>

*  [Illustrate Zero Assumption](makefig_FDReg.html)
*  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
*  [Adaptive Shrinkage](adaptive_shrinkage.html)
*  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
*  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
*  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)

### 2016-05-25 paper/main.tex
add examples

We introduce a novel Empirical Bayes approach for large-scale hypothesis testing, including  
estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
existing approaches to FDR analysis, the method has two key differences. First, it
assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
This ``unimodal assumption" (UA), although natural in many contexts, is very different from assumptions usually made
in FDR analyses, and yields more accurate inferences than existing methods provided that it holds.
The UA also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one 
helps account for variation in measurement precision across tests. It also facilitates the estimation of actual 
effect sizes, and our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
To provide a bridge between interval estimates and significance measures we introduce the term ``local false sign rate"
to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
measure of significance than the local FDR because it is both more generally applicable, and
can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.

@@

Although we have focussed here on applications to FDR, the idea of performing EB inference by flexibly esimating an underlying unimodal prior distribution
could be useful more generally - for example in shrinkage estimation contexts such as wavelet denoising \cite{donoho:1995}. 
Importantly, our work demonstrates that EB inference under a general unimodal assumption is, if anything, computationally simpler  
than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at 
\href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
(The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation.) 
Code and instructions for reproducing analyses and figures in this paper are at
\href{https://github.com/stephenslab/ash}{https://github.com/stephenslab/ash}.

@@

This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, the model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
can allow $g$ to approximate {\it any} unimodal distribution. 
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
This can be relaxed [see (\ref{eqn:beta-alpha})].
\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments are detailed in Detailed Methods.
Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
whatever analysis methods are used \cite{efron2007correlation,leek:2007}.

@@

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$). 
And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

@@

We compare results of \ashr with existing FDR-based methods implemented
in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \href{https://cran.r-project.org/src/contrib/Archive/mixfdr/}{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  
In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.

@@

This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
not limited to this simple simulation. To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
distribution be defined implicitly, by what remains.
 In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
 both methods assume that all $z$ scores near zero (or, equivalently,
all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the 
existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0. See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.

@@

The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
Consequently \ashr will estimate smaller $\lfdr$s and FDRs. 
This is desirable, provided that these estimates remain conservative: that is,
provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behavior. To check its effectiveness
we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
They range from a ``spiky" distribution -- where many non-zero $\beta$ are
too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.

@@

The key point we want to make here is that estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.

@@

Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
if we are prepared to assume that no effects are exactly zero, then removing the point mass 
yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
However, there is ``no free lunch":  if in fact some effects are exactly zero
then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.

@@

Interestingly, even in the ``bimodal" scenario $\ashr$ 
is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
(\href{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}{http://stephenslab.github.io/ash/analysis/check\_mixfdr\_lownoise.html}). 
Of course, we would not advocate the UA in settings where multi-modality is 
clearly visible in the observed $\bhat$.

@@

The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.

@@

With this in place, the penalized log-likelihood for $\pi$ is given by:
\begin{equation}
\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
\end{equation}
where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This is a convex optimization problem, which
can be solved very quickly and reliably using interior point (IP) methods. We used the {\tt KWdual} function from the {\tt R} package {\tt REBayes} \cite{REBayes}, which uses {\tt Rmosek} \cite{Rmosek}. We also found a simple EM algorithm \cite{dempster77}, accelerated using the elegant {\tt R} package {\tt SQUAREM} \cite{varadhan2008simple},
to provide adequate performance. In our EM implementation we initialized $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
and the one-step updates are:
\begin{align} \label{eqn:w}
w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}.
\end{align}
One benefit to the EM algorithm is fewer software dependencies.
Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
See \href{http://stephenslab.github.io/ash/analysis/checkIP.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \href{http://stephenslab.github.io/ash/analysis/IPvsEM.html}{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.

@@

I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
git and github. The {\tt ashr} R package is available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr} and includes contributions 
from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.

>>>>

We introduce a novel Empirical Bayes approach for large-scale hypothesis testing, including  
estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
existing approaches to FDR analysis, the method has two key differences. First, it
assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
This ``unimodal assumption" (UA), although natural in many contexts, is not usually incorporated into standard FDR analysis,
and we demonstrate how incorporating it brings many benefits.  Specifically, the UA
facilitates efficient and robust computation -- estimating the unimodal distribution involves solving a simple convex optimization problem --
and enables more accurate inferences provided that it holds.
Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one 
helps account for variation in measurement precision across tests. It also facilitates the estimation of actual 
effect sizes, and unlike standard FDR analysis methods
our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
To provide a bridge between interval estimates and significance measures we introduce the term ``local false sign rate"
to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
measure of significance than the local FDR because it is both more generally applicable, and
can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
available from \url{http://github.com/stephens999/ashr}.

@@

Although we focus here on FDR applications, the idea of performing EB inference using a flexible unimodal prior distribution
is useful more generally. For example, the methods described here can be applied directly to perform shrinkage estimation for
wavelet denoising \cite{donoho:1995}, an idea explored in a companion paper \cite{xing:2016}. And analogous ideas can be used to perform
EB inference for variances \cite{lu2016variance}.
Importantly, our work demonstrates how EB inference under a general unimodal assumption is, if anything, computationally simpler  
than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.

Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at 
\url{http://github.com/stephens999/ashr}.
(The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation.) 
Code and instructions for reproducing analyses and figures in this paper are at
\url{https://github.com/stephenslab/ash}.

@@

This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, the model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
can allow $g$ to approximate {\it any} unimodal distribution. 
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
This can be relaxed [see (\ref{eqn:beta-alpha})].
\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments, detailed in Detailed Methods, are all implemented in our software.
Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
whatever analysis methods are used \cite{efron2007correlation,leek:2007}.

@@

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$;
see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

@@

We compare results of \ashr with existing FDR-based methods implemented
in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.

@@

This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}). To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
distribution be defined implicitly, by what remains.
 In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
 both methods assume that all $z$ scores near zero (or, equivalently,
all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the 
existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0. See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.

@@

The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
Consequently \ashr will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
This is desirable, provided that these estimates remain conservative: that is,
provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behavior. To check its effectiveness
we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
They range from a ``spiky" distribution -- where many non-zero $\beta$ are
too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.

@@

The key point we want to make here is that estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.

@@

Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
if we are prepared to assume that no effects are exactly zero, then removing the point mass 
yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
However, there is ``no free lunch":  if in fact some effects are exactly zero
then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
We conclude that {\it if} ensuring a ``conservative" analysis is important then one should allow for a point mass at 0.

@@

Interestingly, even in the ``bimodal" scenario $\ashr$ 
is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). 
Of course, we would not advocate the UA in settings where multi-modality is 
clearly visible in the observed $\bhat$.

@@

The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
One general issue is that the methods based on mixtures of uniform distributions often slightly curtail the tail of $g$, causing the probability of very large outlying effects to be understated;  see also \url{http://stephenslab.github.io/ash/analysis/efron.fcr.html}.

@@

With this in place, the penalized log-likelihood for $\pi$ is given by:
\begin{equation}
\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
\end{equation}
where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This is a convex optimization problem, which
can be solved very quickly and reliably using interior point (IP) methods. We used the {\tt KWdual} function from the {\tt R} package {\tt REBayes} \cite{REBayes}, which uses {\tt Rmosek} \cite{Rmosek}. We also found a simple EM algorithm \cite{dempster77}, accelerated using the elegant {\tt R} package {\tt SQUAREM} \cite{varadhan2008simple},
to provide adequate performance. In our EM implementation we initialized $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
and the one-step updates are:
\begin{align} \label{eqn:w}
w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}.
\end{align}
One benefit to the EM algorithm is fewer software dependencies.
Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
See \url{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \url{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.

@@

I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
